\LARGE \bf Stochastic Uncertainty Propagation in Power System Dynamics using Measure-valued Proximal Recursions

### 2022-03-09 Papers/Stochastic_Uncertainty_Propagation_TPS/RebuttalRound1-TPWRS-01391-2021.tex

{\nib {\blue We have made several updates to clarify the technical issues and to improve the presentation.

\ul{Following the comment \textbf{R4.3} to improve the presentation, in the revised manuscript, we have expanded the last two paragraphs of Sec. III-A with a new reference [50]}. Please see our response to that comment below for details.

\ul{As detailed in the response to comment \textbf{R3.4}, in the revised manuscript, we have enlarged Figures 5 and 6, and expanded the plot in Fig. 6}. There we also explained why those plot styles are used.

\ul{Furthermore, Fig. 7 in the revised manuscript has also been expanded to include the univariate angular velocity transient marginal plots for all the five generators of the IEEE 14 bus system; previous version had this plot only for the generator 4 (bus 6)}.

### 2022-08-12 Papers/Stochastic_Uncertainty_Propagation_TPS/JKO_Power_Systems_TPS_Rev2.tex
Rebuttal round 2

 As is well known, the Monte Carlo techniques are easy to apply but the computational cost scales exponentially with the number of dimensions, thus making it prohibitive for realistic power systems dynamic simulation. As an alternative, \emph{probabilistic small signal analysis} \cite{nwankpa1992stochastic,rueda2009assessment,huang2013quasi,dhople2013analysis} have appeared in the power systems literature, albeit at the expense of the additional assumption that the random perturbations remain ``small". \emph{Polynomial chaos and related stochastic collocation methods} \cite{hockenberry2004evaluation,xu2019propagating} can do away with the ``small stochastic perturbation" assumption but due to the finite dimensional approximation of the probability space, computational performance degrades if the long-term statistics are desired. Furthermore, to cope with the stochasticity, these techniques require simulating a higher-dimensional system than the dimension of the physical state space, which further limits the scalability for nonlinear simulation. More recently, approximation methods such as \emph{stochastic averaging} \cite{ju2018analytical,ju2018stochastic} based on certain energy function \cite{pai1989energy,chang1995direct,sauerpai1998} have appeared. In \cite{maldonado2018uncertainty}, an algorithm for \emph{propagating first few statistical moments} was proposed. However, this leads to moment closure problems since the dimensions of the time-varying sufficient statistics associated with the corresponding transient joint state PDFs are not known in general. {\blue{We also note the usage of \emph{stochastic} differential algebraic equations (DAEs) for power system dynamic simulation \cite{wang2011numerical,milano2013systematic}. These studies, however, simulate the sample paths and do not directly propagate the joint state PDFs.}}

@@

For $k\in\mathbb{N}$, and for some chosen step size $h>0$, we discretize time as $t_{k} := kh$, and define the infinite dimensional proximal operator of the functional $h\Phi$ w.r.t. the distance ${\rm{dist}}(\cdot,\cdot)$, given by
\begin{eqnarray}
{\text{prox}}^{{\rm{dist}}}_{h\Phi}(\varrho_{k-1}) := \underset{\varrho\in\mathcal{P}_{2}} {{\rm{arg\:inf}}}\: \frac{1}{2} {\rm{dist}}^{2}\left(\varrho,\varrho_{k-1}\right) + h\: \Phi(\varrho).
\label{InfiniteDimProxOpDef}
\end{eqnarray}
Now consider a proximal recursion over the manifold $\mathcal{P}_{2}$ as
\begin{eqnarray}
\varrho_{k} = {\text{prox}}^{{\rm{dist}}}_{h\Phi}(\varrho_{k-1}), \quad k\in\mathbb{N}, \quad \varrho_{0}(\bm{x}) := \rho_{0}(\bm{x}).
\label{ProxRecursionInfiniteDim}	
\end{eqnarray}
Given the PDE \eqref{KineticFPK}, we would like to design the functional pair $\left(\Phi,{\rm{dist}}\right)$ such that the sequence of functions $\{\varrho_{k}\}_{k\in\mathbb{N}}$ generated by the proximal recursion \eqref{ProxRecursionInfiniteDim}, in the small time step limit, converges to the flow $\rho(t=kh,\bm{\theta},\bm{\omega})$ generated by the PDE initial value problem of interest. In particular\footnote{$L^{1}$ denotes the Lebesgue space of absolutely integrable functions.},
\begin{align}
\varrho_{k}(\bm{\theta},\bm{\omega}) \xrightarrow{h\downarrow 0} \rho(t=kh,\bm{\theta},\bm{\omega}) \;\text{in}\;L^{1}\left(\mathbb{T}^{n} \times \mathbb{R}^{n}\right).
\label{ConvergenceGuarantee}	
\end{align}
We remark here that \eqref{DefP2}, \eqref{InfiniteDimProxOpDef}, \eqref{ProxRecursionInfiniteDim}, \eqref{ConvergenceGuarantee} can be written more generally in {\blue{terms}} of the joint probability measures instead of PDFs, i.e., even when the underlying measures are not absolutely continuous. Fig. \ref{fig:ProxSchematic} illustrates the idea of approximating the joint PDF trajectory $\rho(t,\bm{x})$ through the sequence $\{\varrho_{k}\}_{k\in\mathbb{N}}$ computed from a variational recursion over $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.

@@

The algorithmic appeal of the proximal recursion \eqref{ProxRecursionInfiniteDim} is that it opens up the possibility to compute the solution of the PDE initial value problem via recursive convex minimization. A point cloud-based proximal algorithm was proposed in \cite{caluya2019ACC,caluya2019TAC} which was reported to have very fast runtime. {\blue{The main idea in these references was to \emph{co-evolve} the time-varying state samples (via SDE discretization schemes such as the Euler-Maruyama scheme) as well as the joint PDF values (via \eqref{ProxRecursionInfiniteDim}) evaluated at those samples. The resulting computation is an \emph{online} propagation of the joint PDFs as opposed to the \emph{offline} computation in Monte Carlo or density estimation methods \cite{silverman1998density}. The latter methods only propagate the state samples and then approximate the joint PDFs as post-processing.}}

@@

\subsection{From Anisotropic to Isotropic Degenerate Diffusion}\label{subsec:FromAnisoToIso}
Consider the $2n\times 2n$ matrix
\begin{align}
\bm{\Psi}:= \bm{I}_{2} \otimes \left(\bm{M}\bm{\Sigma}^{-1}\right), \label{DefPsi}	
\end{align}
and define the invertible linear map
\begin{align}
\begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix} \mapsto \begin{pmatrix}
\bm{\xi}\\
\bm{\eta}	
\end{pmatrix} := \bm{\Psi} \begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix}. \label{ChangeOfVar}
\end{align}
Applying It\^{o}'s lemma \cite[Ch. 4.2]{oksendal2013stochastic} to the map \eqref{ChangeOfVar}, and using \eqref{ItoSDEvectorlevel}, we find that the transformed state vector $(\bm{\xi},\bm{\eta})^{\top}$ solves the It\^{o} SDE 
\begin{align}
\begin{pmatrix}
{\mathrm{d}}\bm{\xi}\\
{\mathrm{d}}\bm{\eta}	
\end{pmatrix}
\! = \!\begin{pmatrix}\!
\bm{\eta}\\
-\bm{\Upsilon}\nabla_{\bm{\xi}}U(\bm{\xi}) -\nabla_{\bm{\eta}}F\left(\bm{\eta}\right)  	
\!\end{pmatrix}\!{\mathrm{d}}t + \!\begin{pmatrix}\!
 \bm{0}_{n\times n}\\
 \bm{I}_{n}	
 \!\end{pmatrix}\!{\mathrm{d}}\bm{w},
\label{XiEtaVectorSDE}	
\end{align}
where {\blue{the diagonal matrix $\bm{\Upsilon}:= \left(\prod_{i=1}^{n}\sigma_{i}^{2}m_{i}^{-2}\right)\bm{M\Sigma}^{-2}$ has positive entries along the main diagonal, and}} the potentials
{\small{\begin{subequations}
\begin{align}
U(\bm{\xi}) &:= \left(\displaystyle\sum_{i<j}k_{ij}\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} - \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right) \right.\nonumber\\
&\left.\qquad\qquad-\displaystyle\sum_{i=1}^{n}\!\frac{\sigma_{i}}{m_{i}}P_{i}\xi_{i}\right)\left(\prod_{i=1}^{n}\frac{m_{i}^{2}}{\sigma_{i}^{2}}\right), \label{defU}
%\!-\displaystyle\sum_{i=1}^{n}\!\frac{1}{\sigma_{i}}P_{i}\xi_{i} + \!\!\displaystyle\sum_{i,j=1}^{n}\!\!\dfrac{m_{i}}{\sigma_{i}^{2}}k_{ij}\!\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} - \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right)
\\
F(\bm{\eta}) &:= \dfrac{1}{2}\langle\bm{\eta},\bm{\Sigma}^{-1}\bm{\Gamma}\bm{\eta}\rangle. \label{defF}	
\end{align}
\label{NewPotentials}	
\end{subequations}}}

@@

\subsection{Proximal Update}\label{subsec:ProxUpdate}
Looking at \eqref{NewPotentials} and \eqref{NewPDE}, it is natural to consider an energy functional of the form
\begin{align}
\Phi(\tilde{\rho}) &:= \!\!\displaystyle\int_{{\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)}}\times\mathbb{R}^{n}}\!\!\!\left(\!U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2}\log\tilde{\rho}\!\right)\nonumber\\
&\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta},
\label{LyapButNotProx}	
\end{align}
which is the sum of a potential energy (expected value of $U$), a weighted kinetic energy (expected value of $F$), and an internal energy (scaled negative entropy, the entropy being $-\int\tilde{\rho}\log\tilde{\rho}$). %Indeed, it can be shown that (Appendix \ref{AppLyap}) the functional \eqref{LyapButNotProx} is a Lyapunov-like functional, i.e., $\Phi$ is decreasing along the solution of \eqref{NewPDE}.
However, unlike the gradient drift case mentioned in Sec. \ref{subsec:GenGradDescent}, it is not possible to express the right hand side of \eqref{NewPDE} as the Wasserstein gradient of the functional \eqref{LyapButNotProx}. This can be verified via direct computation by recalling that the Wasserstein gradient is defined as \cite[Ch. 8]{ambrosio2008gradient} 
$$\nabla^{\text{Wasserstein}}\Phi := -\nabla\cdot\left(\rho\nabla\dfrac{\delta\Phi}{\delta\rho}\right),$$
where $\nabla$ denotes the standard Euclidean gradient w.r.t. the vector $(\bm{\xi},\bm{\eta})^{\top}$, and $\frac{\delta}{\delta\rho}$ denotes the functional derivative. %From \eqref{FunctionalDerivative}, \eqref{defvectorfieldv} and \eqref{firstequality} in Appendix A, it is clear that the vector field $\bm{v}$ defined in \eqref{defvectorfieldv} is not equal to the negative of $\nabla\dfrac{\delta\Phi}{\delta\rho}$. 
Thus, 
%even though the functional \eqref{LyapButNotProx} decreases along the transient solution of \eqref{NewPDE}, 
we cannot interpret the flow generated by \eqref{TransformedPDEivp} as the Wasserstein gradient flow of the functional \eqref{LyapButNotProx}. Consequently, in \eqref{InfiniteDimProxOpDef}, we cannot construct $(\Phi,{\rm{dist}})$ by pairing \eqref{LyapButNotProx} with the Wasserstein metric.

To set up a variational recursion of the form \eqref{InfiniteDimProxOpDef} for \eqref{TransformedPDEivp}, we set $(\Phi,{\rm{dist}}) \equiv \left(\widetilde{\Phi},\widetilde{W}_{h,\bm{\Upsilon}}\right)$ where
\begin{align}
\widetilde{\Phi}(\tilde{\rho}) := \!\!\int_{{\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)}} \times \mathbb{R}^{n}}\!\!\left(\!F(\boldsymbol{\eta})+\frac{1}{2} \log \tilde{\rho}\!\right) \tilde{\rho}\: \mathrm{d} \boldsymbol{\xi} \mathrm{d} \boldsymbol{\eta},
\label{Phitilde}	
\end{align}
and 
\begin{align}
&\widetilde{W}_{h,\bm{\Upsilon}}^{2}\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right) := \!\underset{\pi\in\Pi\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right)}{\inf}\!\displaystyle\int_{{\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)^{2}}} \times \mathbb{R}^{2n}}\nonumber\\
&\qquad\qquad\qquad\qquad s_{h,\bm{\Upsilon}}\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right)\differential\pi\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right),
\label{Wtilde}    
\end{align}
wherein $h>0$ denotes the step-size in proximal recursion, and $\Pi\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right)$ denotes the set of joint probability measures over the product space ${\blue{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)^{2}}} \times \mathbb{R}^{2n}$ that have finite second moments  with the first marginal $\tilde{\varrho}$, and the second marginal $\tilde{\varrho}_{k-1}$. The ``ground cost" $s_{h,\bm{\Upsilon}}$ in \eqref{Wtilde} is given by
 \begin{align}
     &s_{h,\bm{\Upsilon}}\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right):=\nonumber\\
    &\bigg\langle\!\!\left(\bar{\bm{\eta}}-\bm{\eta}+h\bm{\Upsilon} \nabla_{\bm{\xi}} \bm{U}(\bm{\xi})\right),\bm{\Upsilon}^{-1}\!\left(\bar{\bm{\eta}}-\bm{\eta}+h\bm{\Upsilon} \nabla_{\bm{\xi}}\bm{U}(\bm{\xi})\right)\!\!\bigg\rangle\nonumber\\
    & + 12 \bigg\langle\!\left(\!\frac{\bar{\bm{\xi}}-\bm{\xi}}{h}-\frac{\bar{\bm{\eta}}-\bm{\eta}}{h}\!\right),\bm{\Upsilon}^{-1}\!\left(\!\frac{\bar{\bm{\xi}}-\bm{\xi}}{h}-\frac{\bar{\bm{\eta}}-\bm{\eta}}{h}\!\right)\!\bigg\rangle.
    \end{align}

@@

That the sequence of functions $\{\tilde{\varrho}_{k}\}$ for $k\in\mathbb{N}$, generated by the proximal recursion
\begin{align}
\tilde{\varrho}_{k} &= {\text{prox}}^{\widetilde{W}}_{h\widetilde{\Phi}}(\tilde{\varrho}_{k-1}) \nonumber\\
&\equiv \underset{\tilde{\varrho}\in\mathcal{P}_{2}}{\arg\inf}\:\frac{1}{2}\widetilde{W}^{2}\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right) + h\:\widetilde{\Phi}(\tilde{\varrho}), \quad \tilde{\varrho}_{0} := \tilde{\rho}_{0},
\label{KineticProx}	
\end{align}
converges to the flow generated by \eqref{TransformedPDEivp}, i.e., 
$$\tilde{\varrho}_{k}(\bm{\xi},\bm{\eta}) \xrightarrow{h\downarrow 0} {\blue{\tilde{\rho}}}(t=kh,\bm{\xi},\bm{\eta})\;\text{in}\;L^{1}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right),$$
can be established following the arguments in \cite{duong2014conservative}. To numerically perform the recursion \eqref{KineticProx}, we employ the proximal algorithm proposed in \cite{caluya2019TAC} with finite number of samples, as explained next.

>>>>

 As is well known, the Monte Carlo techniques are easy to apply but the computational cost scales exponentially with the number of dimensions, thus making it prohibitive for realistic power systems dynamic simulation. As an alternative, \emph{probabilistic small signal analysis} \cite{nwankpa1992stochastic,rueda2009assessment,huang2013quasi,dhople2013analysis} have appeared in the power systems literature, albeit at the expense of the additional assumption that the random perturbations remain ``small". \emph{Polynomial chaos and related stochastic collocation methods} \cite{hockenberry2004evaluation,xu2019propagating} can do away with the ``small stochastic perturbation" assumption but due to the finite dimensional approximation of the probability space, computational performance degrades if the long-term statistics are desired. Furthermore, to cope with the stochasticity, these techniques require simulating a higher-dimensional system than the dimension of the physical state space, which further limits the scalability for nonlinear simulation. More recently, approximation methods such as \emph{stochastic averaging} \cite{ju2018analytical,ju2018stochastic} based on certain energy function \cite{pai1989energy,chang1995direct,sauerpai1998} have appeared. In \cite{maldonado2018uncertainty}, an algorithm for \emph{propagating first few statistical moments} was proposed. However, this leads to moment closure problems since the dimensions of the time-varying sufficient statistics associated with the corresponding transient joint state PDFs are not known in general. We also note the usage of \emph{stochastic} differential algebraic equations (DAEs) for power system dynamic simulation \cite{wang2011numerical,milano2013systematic}. These studies, however, simulate the sample paths and do not directly propagate the joint state PDFs.

@@

For $k\in\mathbb{N}$, and for some chosen step size $h>0$, we discretize time as $t_{k} := kh$, and define the infinite dimensional proximal operator of the functional $h\Phi$ w.r.t. the distance ${\rm{dist}}(\cdot,\cdot)$, given by
\begin{eqnarray}
{\text{prox}}^{{\rm{dist}}}_{h\Phi}(\varrho_{k-1}) := \underset{\varrho\in\mathcal{P}_{2}} {{\rm{arg\:inf}}}\: \frac{1}{2} {\rm{dist}}^{2}\left(\varrho,\varrho_{k-1}\right) + h\: \Phi(\varrho).
\label{InfiniteDimProxOpDef}
\end{eqnarray}
Now consider a proximal recursion over the manifold $\mathcal{P}_{2}$ as
\begin{eqnarray}
\varrho_{k} = {\text{prox}}^{{\rm{dist}}}_{h\Phi}(\varrho_{k-1}), \quad k\in\mathbb{N}, \quad \varrho_{0}(\bm{x}) := \rho_{0}(\bm{x}).
\label{ProxRecursionInfiniteDim}	
\end{eqnarray}
Given the PDE \eqref{KineticFPK}, we would like to design the functional pair $\left(\Phi,{\rm{dist}}\right)$ such that the sequence of functions $\{\varrho_{k}\}_{k\in\mathbb{N}}$ generated by the proximal recursion \eqref{ProxRecursionInfiniteDim}, in the small time step limit, converges to the flow $\rho(t=kh,\bm{\theta},\bm{\omega})$ generated by the PDE initial value problem of interest. In particular\footnote{$L^{1}$ denotes the Lebesgue space of absolutely integrable functions.},
\begin{align}
\varrho_{k}(\bm{\theta},\bm{\omega}) \xrightarrow{h\downarrow 0} \rho(t=kh,\bm{\theta},\bm{\omega}) \;\text{in}\;L^{1}\left(\mathbb{T}^{n} \times \mathbb{R}^{n}\right).
\label{ConvergenceGuarantee}	
\end{align}
We remark here that \eqref{DefP2}, \eqref{InfiniteDimProxOpDef}, \eqref{ProxRecursionInfiniteDim}, \eqref{ConvergenceGuarantee} can be written more generally in terms of the joint probability measures instead of PDFs, i.e., even when the underlying measures are not absolutely continuous. Fig. \ref{fig:ProxSchematic} illustrates the idea of approximating the joint PDF trajectory $\rho(t,\bm{x})$ through the sequence $\{\varrho_{k}\}_{k\in\mathbb{N}}$ computed from a variational recursion over $\mathcal{P}_{2}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right)$.

@@

The algorithmic appeal of the proximal recursion \eqref{ProxRecursionInfiniteDim} is that it opens up the possibility to compute the solution of the PDE initial value problem via recursive convex minimization. A point cloud-based proximal algorithm was proposed in \cite{caluya2019ACC,caluya2019TAC} which was reported to have very fast runtime. The main idea in these references was to \emph{co-evolve} the time-varying state samples (via SDE discretization schemes such as the Euler-Maruyama scheme) as well as the joint PDF values (via \eqref{ProxRecursionInfiniteDim}) evaluated at those samples. The resulting computation is an \emph{online} propagation of the joint PDFs as opposed to the \emph{offline} computation in Monte Carlo or density estimation methods \cite{silverman1998density}. The latter methods only propagate the state samples and then approximate the joint PDFs as post-processing.

@@

\subsection{From Anisotropic to Isotropic Degenerate Diffusion}\label{subsec:FromAnisoToIso}
Consider the $2n\times 2n$ matrix
\begin{align}
\bm{\Psi}:= \bm{I}_{2} \otimes \left(\bm{M}\bm{\Sigma}^{-1}\right), \label{DefPsi}	
\end{align}
and define the invertible linear map
\begin{align}
\begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix} \mapsto \begin{pmatrix}
\bm{\xi}\\
\bm{\eta}	
\end{pmatrix} := \bm{\Psi} \begin{pmatrix}
\bm{\theta}\\
\bm{\omega}	
\end{pmatrix}. \label{ChangeOfVar}
\end{align}
Applying It\^{o}'s lemma \cite[Ch. 4.2]{oksendal2013stochastic} to the map \eqref{ChangeOfVar}, and using \eqref{ItoSDEvectorlevel}, we find that the transformed state vector $(\bm{\xi},\bm{\eta})^{\top}$ solves the It\^{o} SDE 
\begin{align}
\begin{pmatrix}
{\mathrm{d}}\bm{\xi}\\
{\mathrm{d}}\bm{\eta}	
\end{pmatrix}
\! = \!\begin{pmatrix}\!
\bm{\eta}\\
-\bm{\Upsilon}\nabla_{\bm{\xi}}U(\bm{\xi}) -\nabla_{\bm{\eta}}F\left(\bm{\eta}\right)  	
\!\end{pmatrix}\!{\mathrm{d}}t + \!\begin{pmatrix}\!
 \bm{0}_{n\times n}\\
 \bm{I}_{n}	
 \!\end{pmatrix}\!{\mathrm{d}}\bm{w},
\label{XiEtaVectorSDE}	
\end{align}
where the diagonal matrix $\bm{\Upsilon}:= \left(\prod_{i=1}^{n}\sigma_{i}^{2}m_{i}^{-2}\right)\bm{M\Sigma}^{-2}$ has positive entries along the main diagonal, and the potentials
{\small{\begin{subequations}
\begin{align}
U(\bm{\xi}) &:= \left(\displaystyle\sum_{i<j}k_{ij}\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} - \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right) \right.\nonumber\\
&\left.\qquad\qquad-\displaystyle\sum_{i=1}^{n}\!\frac{\sigma_{i}}{m_{i}}P_{i}\xi_{i}\right)\left(\prod_{i=1}^{n}\frac{m_{i}^{2}}{\sigma_{i}^{2}}\right), \label{defU}
%\!-\displaystyle\sum_{i=1}^{n}\!\frac{1}{\sigma_{i}}P_{i}\xi_{i} + \!\!\displaystyle\sum_{i,j=1}^{n}\!\!\dfrac{m_{i}}{\sigma_{i}^{2}}k_{ij}\!\left(\!1 \!- \!\cos\!\left(\!\dfrac{\sigma_{i}}{m_{i}}\xi_{i} - \dfrac{\sigma_{j}}{m_{j}}\xi_{j} - \varphi_{ij}\!\!\right)\!\right)
\\
F(\bm{\eta}) &:= \dfrac{1}{2}\langle\bm{\eta},\bm{\Sigma}^{-1}\bm{\Gamma}\bm{\eta}\rangle. \label{defF}	
\end{align}
\label{NewPotentials}	
\end{subequations}}}

@@

\subsection{Proximal Update}\label{subsec:ProxUpdate}
Looking at \eqref{NewPotentials} and \eqref{NewPDE}, it is natural to consider an energy functional of the form
\begin{align}
\Phi(\tilde{\rho}) &:= \!\!\displaystyle\int_{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)\times\mathbb{R}^{n}}\!\!\!\left(\!U(\bm{\xi})+ F\left(\bm{\eta}\right) + \frac{1}{2}\log\tilde{\rho}\!\right)\nonumber\\
&\qquad\qquad\qquad\qquad\qquad\qquad\qquad\qquad\tilde{\rho}\:\differential\bm{\xi}\:\differential\bm{\eta},
\label{LyapButNotProx}	
\end{align}
which is the sum of a potential energy (expected value of $U$), a weighted kinetic energy (expected value of $F$), and an internal energy (scaled negative entropy, the entropy being $-\int\tilde{\rho}\log\tilde{\rho}$). %Indeed, it can be shown that (Appendix \ref{AppLyap}) the functional \eqref{LyapButNotProx} is a Lyapunov-like functional, i.e., $\Phi$ is decreasing along the solution of \eqref{NewPDE}.
However, unlike the gradient drift case mentioned in Sec. \ref{subsec:GenGradDescent}, it is not possible to express the right hand side of \eqref{NewPDE} as the Wasserstein gradient of the functional \eqref{LyapButNotProx}. This can be verified via direct computation by recalling that the Wasserstein gradient is defined as \cite[Ch. 8]{ambrosio2008gradient} 
$$\nabla^{\text{Wasserstein}}\Phi := -\nabla\cdot\left(\rho\nabla\dfrac{\delta\Phi}{\delta\rho}\right),$$
where $\nabla$ denotes the standard Euclidean gradient w.r.t. the vector $(\bm{\xi},\bm{\eta})^{\top}$, and $\frac{\delta}{\delta\rho}$ denotes the functional derivative. %From \eqref{FunctionalDerivative}, \eqref{defvectorfieldv} and \eqref{firstequality} in Appendix A, it is clear that the vector field $\bm{v}$ defined in \eqref{defvectorfieldv} is not equal to the negative of $\nabla\dfrac{\delta\Phi}{\delta\rho}$. 
Thus, 
%even though the functional \eqref{LyapButNotProx} decreases along the transient solution of \eqref{NewPDE}, 
we cannot interpret the flow generated by \eqref{TransformedPDEivp} as the Wasserstein gradient flow of the functional \eqref{LyapButNotProx}. Consequently, in \eqref{InfiniteDimProxOpDef}, we cannot construct $(\Phi,{\rm{dist}})$ by pairing \eqref{LyapButNotProx} with the Wasserstein metric.

To set up a variational recursion of the form \eqref{InfiniteDimProxOpDef} for \eqref{TransformedPDEivp}, we set $(\Phi,{\rm{dist}}) \equiv \left(\widetilde{\Phi},\widetilde{W}_{h,\bm{\Upsilon}}\right)$ where
\begin{align}
\widetilde{\Phi}(\tilde{\rho}) := \!\!\int_{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right) \times \mathbb{R}^{n}}\!\!\left(\!F(\boldsymbol{\eta})+\frac{1}{2} \log \tilde{\rho}\!\right) \tilde{\rho}\: \mathrm{d} \boldsymbol{\xi} \mathrm{d} \boldsymbol{\eta},
\label{Phitilde}	
\end{align}
and 
\begin{align}
&\widetilde{W}_{h,\bm{\Upsilon}}^{2}\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right) := \!\underset{\pi\in\Pi\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right)}{\inf}\!\displaystyle\int_{\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)^{2} \times \mathbb{R}^{2n}}\nonumber\\
&\qquad\qquad\qquad\qquad s_{h,\bm{\Upsilon}}\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right)\differential\pi\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right),
\label{Wtilde}    
\end{align}
wherein $h>0$ denotes the step-size in proximal recursion, and $\Pi\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right)$ denotes the set of joint probability measures over the product space $\left(\prod_{i=1}^{n}[0,2\pi m_{i}/\sigma_{i})\right)^{2} \times \mathbb{R}^{2n}$ that have finite second moments  with the first marginal $\tilde{\varrho}$, and the second marginal $\tilde{\varrho}_{k-1}$. The ``ground cost" $s_{h,\bm{\Upsilon}}$ in \eqref{Wtilde} is given by
 \begin{align}
     &s_{h,\bm{\Upsilon}}\left(\bm{\xi},\bm{\eta},\bar{\bm{\xi}},\bar{\bm{\eta}}\right):=\nonumber\\
    &\bigg\langle\!\!\left(\bar{\bm{\eta}}-\bm{\eta}+h\bm{\Upsilon} \nabla_{\bm{\xi}} \bm{U}(\bm{\xi})\right),\bm{\Upsilon}^{-1}\!\left(\bar{\bm{\eta}}-\bm{\eta}+h\bm{\Upsilon} \nabla_{\bm{\xi}}\bm{U}(\bm{\xi})\right)\!\!\bigg\rangle\nonumber\\
    & + 12 \bigg\langle\!\left(\!\frac{\bar{\bm{\xi}}-\bm{\xi}}{h}-\frac{\bar{\bm{\eta}}-\bm{\eta}}{h}\!\right),\bm{\Upsilon}^{-1}\!\left(\!\frac{\bar{\bm{\xi}}-\bm{\xi}}{h}-\frac{\bar{\bm{\eta}}-\bm{\eta}}{h}\!\right)\!\bigg\rangle.
    \end{align}

@@

That the sequence of functions $\{\tilde{\varrho}_{k}\}$ for $k\in\mathbb{N}$, generated by the proximal recursion
\begin{align}
\tilde{\varrho}_{k} &= {\text{prox}}^{\widetilde{W}}_{h\widetilde{\Phi}}(\tilde{\varrho}_{k-1}) \nonumber\\
&\equiv \underset{\tilde{\varrho}\in\mathcal{P}_{2}}{\arg\inf}\:\frac{1}{2}\widetilde{W}^{2}\left(\tilde{\varrho},\tilde{\varrho}_{k-1}\right) + h\:\widetilde{\Phi}(\tilde{\varrho}), \quad \tilde{\varrho}_{0} := \tilde{\rho}_{0},
\label{KineticProx}	
\end{align}
converges to the flow generated by \eqref{TransformedPDEivp}, i.e., 
$$\tilde{\varrho}_{k}(\bm{\xi},\bm{\eta}) \xrightarrow{h\downarrow 0} \tilde{\rho}(t=kh,\bm{\xi},\bm{\eta})\;\text{in}\;L^{1}\left(\mathbb{T}^{n}\times\mathbb{R}^{n}\right),$$
can be established following the arguments in \cite{duong2014conservative}. To numerically perform the recursion \eqref{KineticProx}, we employ the proximal algorithm proposed in \cite{caluya2019TAC} with finite number of samples, as explained next.
